Beep Beep!

### 2015-03-10 FirstDraftMaterials/Draft.tex

% PW Model: Hugoniot Loci and Integral Curves
\subsection{Hugoniot Loci and Integral Curves}
Brisa

@@

\subsubsection{Hugoniot Loci}
In order to consider the Hugoniot loci for this system we must first write the system in conversation form. 
To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
by $\rho$ gives us the two equations 
\begin{align*}
&(v + p(\rho ))\partial_t\rho + (v + p(\rho ))\partial_x(\rho v) = 0,  \\
&\rho\partial_t \left(v + p(\rho )\right) + \rho v\partial_x \left( v + p(\rho )\right) = 0.
\end{align*}
Adding these two equations gives
\begin{align}\label{AR:eq2}
\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align}
Then from (\ref{AR:eq1}) and (\ref{AR:eq2}) we have the system 
\begin{align*}
&\partial_t\rho + \partial_x(\rho v) = 0, \\
&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align*}
This can be rewritten as $q_t + f(q)_x = $ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ y
\end{matrix}\right], \hspace{0.3in}
f(q) = \left[ \begin{matrix}
v\rho \\
vy
\end{matrix}\right].
\end{align*}
where $y = \rho\left(v + p(\rho )\right)$.
The Rankine-Hugoniot condition tells us that 
\begin{align*}
s(q_*- q) = f(q_*) - f(q).
\end{align*}
Therefore, from this condition we get the two equations
\begin{align}
s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AR:eq3}\\
s(y_* - y) = v_*y_* - vy.\label{AR:eq4}
\end{align}
With a little rearranging and by plugging in for $y$, these two equations can be rewritten as
\begin{align}
&\rho_*(s - v_*) = \rho (s - v), \label{AR:eq5}\\
&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AR:eq6}
\end{align}
From (\ref{AR:eq5}), we know that there are two cases:

(a) $\rho_*(s - v_*) = \rho (s - v) = 0$. In this case we have that
\begin{align*}
s - v_* = s - v 
\end{align*}
except if one of the two densities, $\rho_*$, $\rho$ is zero. WHY DOES THIS NOT MATTER? Therefore, we can say that
\begin{align*}
v_* = v.
\end{align*}
Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AR:eq4}) gives
\begin{align*}
y_*(s - v) = y(s - v).
\end{align*}
So, either $y = y_*$ or $s = v$. If $y = y_*$ we have a stationary point in the $Y = (\rho, y)$ plane, which does not 
describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
Therefore this loci in the $Y$ plane is given by 
\begin{align*}
y = \rho ( v_* + p(\rho )).
\end{align*}
(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
\begin{align*}
\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
v_* + p(\rho_* )&= v + p(\rho ).
\end{align*}
Therefore, the Hugoniot loci for this case in the Y plane are given by
\begin{align*}
y = v_* + p(\rho_* ),
\end{align*}
since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
\begin{align*}
s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
\end{align*}
Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
\begin{align*}
s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
\approx v - \rho p'(\rho)
\end{align*}
for $\rho_* \approx \rho$.

@@

@@

\begin{figure}[h!]
 \centering
 \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U.jpg}
   }
    \subfigure[Valid waves for the point $q_l$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_ql.jpg}
   }
    \subfigure[Valid waves for the point $q_r$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_qr.jpg}
   }
    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
   }
 \subfigure[Valid waves for the point $q_l$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
   }
 \subfigure[Valid waves for the point $q_r$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
   }
 \caption[Optional caption for list of figures]
 {Figures showing the calculation of the valid waves for a given point. Figures (a)
  and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
  in magenta. The other figures show valid shocks in red and valid rarefaction 
  waves in green. The contact discontinuity is shown in blue.}
   \label{fig:AR_validity}
\end{figure}

@@

\begin{figure}[h!]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
   }
% TODO: add subfigure for example1 in PW
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 1 from the AR and PW models.}
  \label{fig:example1}
\end{figure}

\begin{figure}[h!]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
   }
% TODO: add subfigure for example2 in PW
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 2 from the AR and PW models.}
  \label{fig:example2}
\end{figure}

\begin{table}[t]
\caption{Initial values used in examples 1 and 2.}
\begin{center}
\begin{tabular}{| c | c c |}
\hline
& Example 1 & Example 2\\
\hline
$v_l$ & 0.2 & 0 \\
$\rho_l$ & 0.3 & 0\\
$v_r$ & 0.3 & 0 .3\\
$\rho_r$ & 0.4 & 0.4\\
\hline
\end{tabular}
\end{center}
 \label{table:1}
\end{table}

Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}. Note that in example 1, shown in Figure \ref{fig:example1}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity. The middle state is denoted by $\tilde{m}$ in the figure. In example 2, show in Figure \ref{fig:example2}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is a rarefaction wave, and there is no middle state.

>>>>

% PW Model: Hugoniot Loci and Integral Curves
\subsection{Hugoniot Loci and Integral Curves}
In \cite{AwRascle2000}, they consider the simplified PW model
\begin{align*}
&\rho_t + \left( \rho v\right)_x = 0 \\
&\left( \rho v\right)_t + \left( \rho v^2 + p(\rho )\right)_x = 0,
\end{align*}
which is equivalent to the system shown above if we let $1/\tau = 0$ and $\mu = 0$.
From this system, the equation for the Hugoniot loci can be found to be
\begin{align*}
\rho v = \rho v_* \pm \rho \sqrt{\left( \rho - \rho_*\right) \left( p(\rho ) - p(\rho_*)\right) / \rho_*\rho }.
\end{align*}
Using $p(\rho) = \rho^{\gamma}$ for $\gamma \neq 1$, since this is the value used in the AR model we will examine later, the equations for the integral curves can be found to be
\begin{align*}
\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho_*)} - \sqrt{p'(\rho)}\right)
\end{align*}
for $\lambda_1$, and 
\begin{align*}
\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho)} - \sqrt{p'(\rho_*)} \right)
\end{align*}
for $\lambda_2$.

\begin{figure}[h!]
 \centering
 \subfigure[Hugonio loci for $\lambda_1$ and $\lambda_2$ shown in magenta and red, respectively.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Loci.jpg}
   }
 \subfigure[Integral curves for $\lambda_1$, $\lambda_2$ shown in light and dark green, respectively.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_IntegralCurves.jpg}
   }
 \subfigure[$\lambda_1$ loci and integral curve passing through
  a point $(\rho_*, v_*)$, with $\lambda_1$ contour lines shown in cyan.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Lambda1Curves.jpg}
   }
    \subfigure[$\lambda_2$ loci and integral curve passing through
  a point $(\rho_*, v_*)$, with $\lambda_2$ contour lines shown in cyan.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Lambda2Curves.jpg}
   }
 \subfigure[Valid curves for a point $q_l$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesLeft.jpg}
   }
 \subfigure[Valid curves for a point $q_r$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesRight.jpg}
   }
 \caption[Optional caption for list of figures]
 {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$ in the $M = (\rho, \rho v)$ plane. Note that loci are shown in red and magenta and integral curves are shown in green.}
  \label{fig:PW_curves}
\end{figure}

Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point. When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease. When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase. In Figure \ref{fig:PW_curves} (c) and (d) we see the Hugoniot loci and integral curves for this system with the contour lines for $\lambda_1$ and $\lambda_2$. Using these plots, we can determine which areas of the curves are valid shocks and which are valid rarefaction waves. Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed. Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.

@@

\subsubsection{Hugoniot Loci}
In order to consider the Hugoniot loci for this system we must first write the system in conservation form. 
To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
by $\rho$ gives us the two equations 
\begin{align*}
&(v + p(\rho ))\partial_t\rho + (v + p(\rho ))\partial_x(\rho v) = 0,  \\
&\rho\partial_t \left(v + p(\rho )\right) + \rho v\partial_x \left( v + p(\rho )\right) = 0.
\end{align*}
Adding these two equations gives
\begin{align}\label{AR:eq2}
\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align}
Then from (\ref{AR:eq1}) and (\ref{AR:eq2}) we have the system 
\begin{align*}
&\partial_t\rho + \partial_x(\rho v) = 0, \\
&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align*}
This can be rewritten as $q_t + f(q)_x = $ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ y
\end{matrix}\right], \hspace{0.3in}
f(q) = \left[ \begin{matrix}
v\rho \\
vy
\end{matrix}\right].
\end{align*}
where $y = \rho\left(v + p(\rho )\right)$.
The Rankine-Hugoniot condition tells us that 
\begin{align*}
s(q_*- q) = f(q_*) - f(q).
\end{align*}
Therefore, from this condition we get the two equations
\begin{align}
s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AR:eq3}\\
s(y_* - y) = v_*y_* - vy.\label{AR:eq4}
\end{align}
With a little rearranging and by plugging in for $y$, these two equations can be rewritten as
\begin{align}
&\rho_*(s - v_*) = \rho (s - v), \label{AR:eq5}\\
&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AR:eq6}
\end{align}
From (\ref{AR:eq5}), we know that there are two cases:

(a) $\rho_*(s - v_*) = \rho (s - v) = 0$. In this case we have that
\begin{align*}
s - v_* = s - v 
\end{align*}
except if one of the two densities, $\rho_*$, $\rho$ is zero. Therefore, we can say that
\begin{align*}
v_* = v.
\end{align*}
Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AR:eq4}) gives
\begin{align*}
y_*(s - v) = y(s - v).
\end{align*}
So, either $y = y_*$ or $s = v$. If $y = y_*$ we have a stationary point in the $Y = (\rho, y)$ plane, which does not 
describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
Therefore this loci in the $Y$ plane is given by 
\begin{align*}
y = \rho ( v_* + p(\rho )).
\end{align*}

(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
\begin{align*}
\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
v_* + p(\rho_* )&= v + p(\rho ).
\end{align*}
Therefore, the Hugoniot loci for this case in the Y plane are given by
\begin{align*}
y = v_* + p(\rho_* ),
\end{align*}
since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
\begin{align*}
s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
\end{align*}
Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
\begin{align*}
s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
\approx v - \rho p'(\rho)
\end{align*}
for $\rho_* \approx \rho$.

@@

\begin{figure}[h!]
 \centering
 \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U.jpg}
   }
    \subfigure[Valid waves for the point $q_l$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_ql.jpg}
   }
    \subfigure[Valid waves for the point $q_r$ in the U plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_qr.jpg}
   }
    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
   }
 \subfigure[Valid waves for the point $q_l$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
   }
 \subfigure[Valid waves for the point $q_r$ in the M plane.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
   }
 \caption[Optional caption for list of figures]
 {Figures showing the calculation of the valid waves for a given point. Figures (a)
  and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
  in cyan. The other figures show valid shocks in red and valid rarefaction 
  waves in green. The contact discontinuity is shown in blue.}
   \label{fig:AR_validity}
\end{figure}

@@

@@

\begin{figure}[h!]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
   }
    \subfigure[Solution from the PW model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 1 from the AR and PW models.}
  \label{fig:example1}
\end{figure}

\begin{figure}[h!]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
   }
 \subfigure[Solution from the PW model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 2 from the AR and PW models.}
  \label{fig:example2}
\end{figure}

\begin{table}[t]
\caption{Initial values used in examples 1 and 2.}
\begin{center}
\begin{tabular}{| c | c c  c c|}
\hline
& $v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
\hline
Example 1 & 0.2 & .03 & 0.3 & 0.4 \\
Example 2 & 0 & 0 & 0.3 & 0.4\\
\hline
\end{tabular}
\end{center}
 \label{table:1}
\end{table}

Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}. Note that in the solution from the AR model for example 1, shown in Figure \ref{fig:example1} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity. The middle state is denoted by $q_m$ in the figure. In the solution given by the AR model for example 2, show in Figure \ref{fig:example2} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is a shock wave, and there is no middle state.

### 2015-03-10 Presentation/presentation.tex

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Introduction} 
Kelsey
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

\begin{frame}
\frametitle{LWR}
Kelsey
\end{frame}

\begin{frame}
\frametitle{PW}
Intro, Kelsey. Eigenvectors and values
\end{frame}

@@

\begin{frame}
\frametitle{AR}
Intro, Kelsey. Eigenvectors and values
\end{frame}

>>>>

\begin{frame}
\titlepage % Print the title page as the first slide
\includegraphics[width=4.75in]{mariokart.png}
\end{frame}

\begin{frame}
\frametitle{Introduction} 
\begin{itemize}
\item Are second-order models always better than first-order models?
\item Not always! The PW model for traffic flow, meant to improve the LWR model's behavior near shocks, actually performs worse than the original in some cases.
\item The PW model is based on compressible fluid dynamics, which ignores some fundamental differences between fluids and cars. Because of this, it can sometimes predict negative velocities.
\item The AR model corrects the problem of negative velocities by utilizing a convective derivative.
\end{itemize}
\end{frame}

%----------------------------------------------------------------------------------------
%	LWR Model
%----------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Lighthill-Whitham-Richards (LWR) Model}

Proposed by Lighthill and Whitham (1955) and Richards (1956), the LWR model describes traffic flow on a single one-way road without entrances or exits:

\[ \rho_t + \Big(\rho V(\rho)\Big)_x = 0 \]

\begin{itemize}
\item $\rho$ = density
\item $V(\rho)$  = preferred velocity, a given nonincreasing function of $\rho$, nonnegative for $\rho$ between 0 and $\rho_m$ (the``jam" density)
\item Predicts piece-wise smooth density, with transitions between regions approximated by shocks
\item Problem: Doesn't adequately describe the motion of cars passing through shocks (cars change velocity instantaneously) 
\end{itemize}

%------------------------------------------------
%	PW Model Background
%------------------------------------------------

\begin{frame}
\frametitle{Payne-Whitham (PW) Model}

Viewing traffic as a compressible fluid, Payne (1971) and Whitham (1974) introduced a second equation analogous to the conservation of momentum, $\rho v$, in fluids:

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] (\rho v)_t + \Big(\rho v^2 + p(\rho)\Big)_x = 0 \end{matrix*} \right. \]

\begin{itemize}
\item $\rho$ = density
\item $v$ = velocity
\item $p(\rho)$ = ``anticipation factor", describes how a driver reacts to variations in density with respect to space
\end{itemize}

\begin{frame}
\frametitle{Payne-Whitham (PW) Model}

Rewriting the second equation in terms of velocity and adding relaxation and viscosity terms, we get the full PW Model:

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{matrix*} \right. \]

\begin{itemize}
\item $\tau$ and $\mu$ are nonnegative constants
\item Relaxation term: represents driver's reaction time
\item Diffusion term: represents the driver's awareness of conditions ahead (and behind)
\end{itemize}

\begin{frame}
\frametitle{Payne-Whitham (PW) Model}

\hspace{2em} A fundamental difference between fluids and traffic: \\[3ex]

\hspace{2em} ``A fluid particle responds to stimuli from the front and from \\ 
\hspace{2em} behind, but a car is an anisotropic particle that mostly \\
\hspace{2em} responds to frontal stimuli." CITATION

\begin{frame}
\frametitle{Payne-Whitham (PW) Model}

\hspace{5em} Homogeneous model: \hspace{7em} Eigenvalues:

\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v\end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = 0 \hspace{3em} \lambda = v \pm \sqrt{p'(\rho)} \] \\[5ex]

From the eigenvalues we see that one characteristic speed is greater than the velocity, which means future conditions are partly determined by conditions behind. This produces an undesirable effect: negative velocities!

@@

%------------------------------------------------
%	AR Model Background
%------------------------------------------------

\begin{frame}
\frametitle{A thought experiment}

Consider the situation where a driver is traveling with speed $v$.  If the density ahead of him is increasing with respect to $x$ but decreasing with respect to $x - vt$, will he speed up or slow down?

\begin{figure}
\centering
\includegraphics[width=2in]{experiment.png}
\end{figure}

\begin{frame}
\frametitle{A thought experiment}

\hspace{2em} Since density ahead is increasing with respect to $x$, the PW \\
\hspace{2em} model predicts that the driver will slow down (as we will see in \\
\hspace{2em} our examples).  However, since the traffic ahead is traveling faster, \\
\hspace{2em} most drivers would actually accelerate.

\begin{frame}
\frametitle{Aw-Rascle (AR) Model}

Aw and Rascle (2000) claim that instead of depending upon the derivative of pressure with respect to $x$, the anticipation factor should involve the convective derivative:

\[ \partial_t + v \, \partial_x \] \\[5ex]

This change reflects the fact that the driver's perspective is a moving frame of reference.

\begin{frame}
\frametitle{Aw-Rascle (AR) Model}

Using the convective derivative, the proposed AR model is: \\[1ex]

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] (v + \rho)_t + v(v + \rho)_x = 0 \end{matrix*} \right. \]

\begin{frame}
\frametitle{Aw-Rascle (AR) Model}

AR model in conservation form: \\[3ex]

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] \Big(\rho (v + p(\rho)) \Big)_t + \Big(\rho v(v + p(\rho)) \Big)_x = 0 \end{matrix*} \right. \] \\[5ex]

Conserved quantities:
\begin{enumerate}
\item Mass (density): $\rho$
\item ``Momentum": $y = v + p(\rho)$
\end{enumerate}

\begin{frame}
\frametitle{Aw-Rascle (AR) Model}

AR model rewritten in terms of density and velocity: \\[1ex]

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] v_t + \Big( v - \rho p'(\rho) \Big)_x = 0 \end{matrix*} \right. \] \\[3ex]

\hspace{5em} Linearized system: \hspace{7em} Eigenvalues:
\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] 0 & v - \rho p'(\rho)  \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = 0 \hspace{5ex} \begin{matrix*}[l] \lambda_1 = v - \rho p'(\rho) \\[1ex] \lambda_2 = v \end{matrix*} \] \\[1ex]

Note that neither characteristic speed is greater than the velocity!

\begin{frame}
\frametitle{Aw-Rascle (AR) Model}

Aw and Rascle (2000) propose some conditions for traffic models:
\begin{enumerate}
\item The system must be hyperbolic.
\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
\item When solving the Riemann problem with data $Q_{\pm} = (\rho_{\pm}, v_{\pm})$, all waves connecting any state $Q = (\rho, v)$ to its left (behind it) must have a propagation speed (eigenvalue or shock speed) at most equal to the velocity $v$.  
\item The solution to the Riemann problem must agree with the qualitative properties that each driver practically observes every day.
\item Near the vacuum, the solution to the Riemann problem must be very sensitive to the data (no continuous dependence with respect to initial data at $\rho = 0$). 
\end{enumerate}
